\documentclass[./analysis.tex]{subfiles}

\begin{document}
    \section{Appendix}
    \subsection*{References}
    \begin{enumerate}
        \item Mathematics 104 at UC Berkeley, Summer 2019, taught by Michael Christianson.
        \item Mathematics 202A at UC Berkeley, Fall 2019, taught by Professor Marc Rieffel.
        \item Real Analysis for Graduate Students, by Richard F. Bass. \url{http://bass.math.uconn.edu/rags010213.pdf}
        \item Elementary Analysis, by Kenneth A. Ross. (Zero is a natural number, change my mind.)
        \item Cauchy's Construction of $\R$, by Todd Kemp. \url{http://www.math.ucsd.edu/~tkemp/140A/Construction.of.R.pdf}
        \item Renzo's Math 490: Introduction to Topology Lecture Notes. \url{https://www.math.colostate.edu/~renzo/teaching/Topology10/Notes.pdf}
        \item Many answers on Math StackExchange, but particularly this complete construction of the reals (including the algebraic approach): \url{https://math.stackexchange.com/questions/11923/completion-of-rational-numbers-via-cauchy-sequences}
        \item \url{https://www.math.unl.edu/~gmeisters1/papers/Measure/measure.pdf}
        \item Stylistically inspired by Evan Chen's An Infinitely Large Napkin: \url{https://venhance.github.io/napkin/Napkin.pdf}
    \end{enumerate}

    \subsection*{Acknowledgements}

    Thanks to Charlie Cummings, Abhinav Kejriwal, Shaunak Modak, Grant Posner, Shamay Samuel, Kanyes Thaker, and Ellen Thompson for their feedback, and to Kanyes Thaker for some \LaTeX \ code (the coloured boxes). Thanks also to Shivani Kishnani for thinking this was a Math 104 textbook.
\end{document}